
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 12: Predictive Modeling - Part 3"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment12.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


1. Use the `PimaIndiansDiabetes` dataset. Use 15% data for testing. Use cross-validation with of 7 folds to tune random forest `(method='ranger')`.  What are the parameters that produce the greatest accuracy? What is the testing accuracy. 
```{r}
library(mlbench)
library(caret)
library(ranger)

data("PimaIndiansDiabetes", package="mlbench")
set.seed(123)

trainIndex <- createDataPartition(PimaIndiansDiabetes$diabetes, p = .85, list = FALSE)
trainData <- PimaIndiansDiabetes[trainIndex,]
testData  <- PimaIndiansDiabetes[-trainIndex,]
rfGrid <- expand.grid(
  mtry        = c(2, 3, 4, 5),
  splitrule   = c("gini"),
  min.node.size = c(1, 3, 5)
)

fitControl <- trainControl(
  method      = "cv",
  number      = 7,
  classProbs  = TRUE
)

rfFit <- train(
  diabetes ~ .,
  data       = trainData,
  method     = "ranger",
  tuneGrid   = rfGrid,
  trControl  = fitControl,
  metric     = "Accuracy"
)

print(rfFit)

```

2. Use the `PimaIndiansDiabetes` dataset. Go to https://topepo.github.io/caret/available-models.html and pick a classification model.  Tune the classification model using cross-validation of 7 folds. 
```{r}
library(mlbench)
library(caret)

data("PimaIndiansDiabetes")
set.seed(123)

trainIndex <- createDataPartition(PimaIndiansDiabetes$diabetes, p = 0.85, list = FALSE)
trainData  <- PimaIndiansDiabetes[trainIndex, ]
testData   <- PimaIndiansDiabetes[-trainIndex, ]

ctrl <- trainControl(
  method = "cv",
  number = 7,
  classProbs = TRUE
)
svmFit <- train(
  diabetes ~ .,
  data = trainData,
  method = "svmRadial",
  trControl = ctrl,
  tuneLength = 10,     # caret automatically expands useful tuning values
  metric = "Accuracy"
)
svmFit
pred <- predict(svmFit, newdata = testData)
confusionMatrix(pred, testData$diabetes)
```

3. (Model Comparison) Use the `PimaIndiansDiabetes` dataset. Pick two models at [this link](https://topepo.github.io/caret/available-models.html) to compare using 7-fold cross validation method. Evaluate the accuracy of the final model on the test data. 
```{r}
set.seed(123)
data("PimaIndiansDiabetes")
trainIndex <- createDataPartition(PimaIndiansDiabetes$diabetes, p = 0.85, list = FALSE)
trainData <- PimaIndiansDiabetes[trainIndex, ]
testData  <- PimaIndiansDiabetes[-trainIndex, ]
ctrl <- trainControl(method = "cv", number = 7, classProbs = TRUE)

knnFit <- train(
  diabetes ~ .,
  data = trainData,
  method = "knn",
  trControl = ctrl,
  tuneLength = 10,
  preProcess = c("center", "scale"),
  metric = "Accuracy"
)

svmFit <- train(
  diabetes ~ .,
  data = trainData,
  method = "svmRadial",
  trControl = ctrl,
  tuneLength = 10,
  preProcess = c("center", "scale"),
  metric = "Accuracy"
)

print(knnFit)
print(svmFit)

knnPred <- predict(knnFit, newdata = testData)
svmPred <- predict(svmFit, newdata = testData)

knnCM <- confusionMatrix(knnPred, testData$diabetes)
svmCM <- confusionMatrix(svmPred, testData$diabetes)

list(
  knn_accuracy = knnCM$overall["Accuracy"],
  svm_accuracy = svmCM$overall["Accuracy"]
)
```

