
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------


-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.   
```{r}
library(readr)
census <- read_csv("adult_census_missing.csv")
colSums(is.na(census))
```

2. Using function `aggr` the VIM package to plot the number of NAs for each column. 
```{r}
library(tidyverse)
library(VIM)
aggr(
  census,
  numbers = TRUE,
  prop = FALSE,
  cex.axis = 0.7,
  gap = 2
)
```

3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 

```{r}
sapply(census, function(x) sum(x == "?", na.rm = TRUE))
sapply(census, function(x) sum(x == "", na.rm = TRUE))
sapply(census, function(x) sum(trimws(x) == "", na.rm = TRUE))
```

4. Replace all the forms of missing values found to NA
```{r}
census[census == "?"] <- NA
census[census == ""] <- NA
census[trimws(census) == ""] <- NA
```

5. Replot the number of NAs for each column. 
```{r}
library(VIM)
aggr(census,
     numbers = TRUE,      
     prop = FALSE,        
     sortVars = TRUE,     
     cex.axis = 0.7,      
     gap = 3,
     ylab = c("Missing Data", "Pattern"))
```

6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
census_noNA <- na.omit(census)
nrow(census_noNA)
```

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
library(tidyr)
census_fill <- census %>%
  fill(everything(), .direction = "down") %>%
  fill(everything(), .direction = "up")
colSums(is.na(census_fill))
```

8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
library(dplyr)

census_medmode <- census %>%
  mutate(across(
    .cols = everything(),
    .fns = ~ if (is.numeric(.x)) {
      replace(.x, is.na(.x), median(.x, na.rm = TRUE))
    } else {
      majority <- names(sort(table(.x), decreasing = TRUE))[1]
      replace(.x, is.na(.x), majority)
    }
  ))
```




