---
title: "Assignment 2"
author: "Andrew Johnson"
output:
  html_document: default
  pdf_document: default
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment2.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Sample Codes***: You can find similar codes for this assignment here [link](assignment2_sample_code.html)


***Submissions***: Submit the link on Github of the assignment to Canvas under Assignment 2. 


Problems
-------
-------


1. Calculate: 


$$
S = 1 + 4 + 7 + 10 +... + 100
$$
```{r}
x <- seq(1,100,3)
sum(x)
```


2. Calculate: 


$$
S = 1^{100} + 2^{100} + 3^{100} +...+100^{100}
$$
```{r}
y <- c(1:100)
sum(y^100)
```

3. Calculate: 

$$
S = \frac{1}{1\cdot 2}+\frac{1}{2 \cdot 3}+...+\frac{1}{100\cdot 101}
$$
```{r}
w <- seq(1,100,1)
z <- seq(2,101,1)
sum(1/(w*z))
```


4. Calculate: 

$$
S = \frac{1}{1\cdot 2 \cdot 3}+\frac{1}{2 \cdot 3 \cdot 4}+...+\frac{1}{100\cdot 101 \cdot 102}
$$
```{r}
a <- seq(1,100,1)
b <- seq(2,101,1)
c <- seq(3,102,1)

sum(1/(a*b*c))
```


5. Calculate: 

$$
S = 1^1 + 3^3 + 5^5 + 7^7 +....+101^{101}
$$
```{r}
t <- seq(1,101,2)
sum(t^t)
```

